Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2016-07-22 3dpas_intro.tex
adding rebuttal document - probably should be turned in .tex later made changes for reviewer 1, and indicated that we have so in the rebuttal

Starting with the experience of the attendees of the 3DPAS workshops,
we have assembled a set of D3 ``applications'' that we describe in
this section.  We realize that we have overloaded the term application
here.  In this document, some of our applications are really {\em
  traditional applications}, meaning a program that is developed by a
user or project to try to answer a science question, usually with some
input data needed and some output data produced.  Applications in this
category are: NGS Analytics (\S\ref{bioSilvia}), CMB
(\S\ref{astroJulian}), Fusion (\S\ref{fusionScott}), Industrial
Incident Notification and Response (\S\ref{Kees}), MODIS Data
Processing (\S\ref{modisKeith}), and Distributed Network Intrusion
Detection (\S\ref{detectionJon}).

Another group of applications are {\em infrastructural applications}.
Applications of this type are collaborative, in that they depend on a
context that is often agreed upon by a community, most often in terms
of how data is stored by the community and how it is accessed, as well
as, more often than not, a weak consensus on the infrastructure to be
used. %\jhanote{Dan: Please confirm you are OK with this modified
%  defintion of infrastructural application}
This allows different
sub-applications (that we again refer to as `applications') to focus
on different stages, such as generating data, or processing data.
Answering a science question now may involve a set of applications
that need to be run in series, perhaps in different stages, that may
be run by different groups that do not frequently interact.  This type
of application includes: ATLAS/WLCG (\S\ref{WLCGSteve}), LSST
(\S\ref{astroAdam}), SOA Astronomy (\S\ref{astroSOAAdam}), Sensor
Network Application (\S\ref{sensorSimon}), Climate
(\S\ref{climateDan}), Interactive Exploration of Environmental Data
(\S\ref{envJon}), and Power Grids (\S\ref{power}).

@@

In \S\ref{sec:distdyndata} we characterized dynamic and distributed
properties of the D3 Science applications. Our analysis was
qualitative by necessity, due to the fact that approaches to quantify
dynamism and distributed properties were often intimately tied to
application/scenario definition and specific usage modes.
Notwithstanding, to quantify better dynamic properties of an
application, we believe a term analogous to the use of Amdahl's
number~\cite{10.1109/ICDE.2000.839382} could be used to characterize `static'
data-intensive applications and systems:

\begin{itemize}
\item[] $\alpha$ = Num. of Compute Operations / Amount of Data
  (Consumed, Generated or Transformed)
\end{itemize}

@@

Similarly, another term can be used to quantify the
``scale'' of distribution, namely the degree of distribution as a
function of the amount of data transferred, where the degree of
distribution can be defined as the number of distinct data locations:

\begin{itemize}
\item[] $\beta$ = Amount of Data Transferred / Degree of
  Distribution
% \item[] $\beta_2$ = Number of Compute Operations / Degree of
%   Distribution
\end{itemize}

@@

It has proven difficult to provide ``global'' quantitative values of
these ratios, as often the value of the numerators is
subject to how an application/scenario is defined, even more so for a
``class of applications'' as opposed to a specific application. We
believe, nonetheless, that these measures suggest a possible way to
quantify an application's dynamism and distribution and thereby find
possible common approaches solutions.  We plan to explore this in our
future work, and also hope others will consider building on these
initial concepts.

>>>>

Starting with the experience of the attendees of the 3DPAS workshops,
we have assembled a set of D3 ``applications'' that we describe in
this section.  We realize that we have overloaded the term application
here.  In this document, some of our applications are really {\em
  traditional applications}, meaning a program that is developed by a
user or project to try to answer a science question, usually with some
input data needed and some output data produced.  Applications in this
category are: NGS (next generation sequencing) Analytics (\S\ref{bioSilvia}), CMB (cosmic microwave background)
(\S\ref{astroJulian}), Fusion (\S\ref{fusionScott}), Industrial
Incident Notification and Response (\S\ref{Kees}), MODIS (moderate resolution imaging spectroradiometer) Data
Processing (\S\ref{modisKeith}), and Distributed Network Intrusion
Detection (\S\ref{detectionJon}).

Another group of applications are {\em infrastructural applications}.
Applications of this type are collaborative, in that they depend on a
context that is often agreed upon by a community, most often in terms
of how data is stored by the community and how it is accessed, as well
as, more often than not, a weak consensus on the infrastructure to be
used. %\jhanote{Dan: Please confirm you are OK with this modified
%  defintion of infrastructural application}
This allows different
sub-applications (that we again refer to as `applications') to focus
on different stages, such as generating data, or processing data.
Answering a science question now may involve a set of applications
that need to be run in series, perhaps in different stages, that may
be run by different groups that do not frequently interact.  This type
of application includes: ATLAS/WLCG (a toroidal large hadron collider apparatus/worldwide large hadron collider computing grid, \S\ref{WLCGSteve}), LSST
(large synoptic survey telescope, \S\ref{astroAdam}), SOA (service-oriented architecture) Astronomy (\S\ref{astroSOAAdam}), Sensor
Network Application (\S\ref{sensorSimon}), Climate
(\S\ref{climateDan}), Interactive Exploration of Environmental Data
(\S\ref{envJon}), and Power Grids (\S\ref{power}).

@@

In \S\ref{sec:distdyndata} we characterized dynamic and distributed
properties of the D3 Science applications. Our analysis was
qualitative by necessity, due to the fact that approaches to quantify
dynamism and distributed properties were often intimately tied to
application/scenario definition and specific usage modes.
Notwithstanding, to quantify better dynamic properties of an
application, we believe a term analogous to the use of Amdahl's
number~\cite{10.1109/ICDE.2000.839382} such as 
\begin{itemize}
\item[] Num. of Compute Operations / Amount of Data
  (Consumed, Generated or Transformed)
\end{itemize}
  could be used to characterize `static'
data-intensive applications and systems:

%\begin{itemize}
%\item[] $\alpha$ = Num. of Compute Operations / Amount of Data
%  (Consumed, Generated or Transformed)
%\end{itemize}

@@

Similarly, another term 
\begin{itemize}
\item[] Amount of Data Transferred / Degree of
  Distribution
\end{itemize}
could be used to quantify the
``scale'' of distribution, namely the degree of distribution as a
function of the amount of data transferred, where the degree of
distribution can be defined as the number of distinct data locations:

%\begin{itemize}
%\item[] $\beta$ = Amount of Data Transferred / Degree of
%  Distribution
%\end{itemize}

@@

It has proven difficult to provide ``global'' quantitative values of
these ratios, as often the value of the numerators is
subject to how an application/scenario is defined, even more so for a
``class of applications'' as opposed to a specific application. We
believe, nonetheless, that these measures (which we refer to as $\alpha$ and $\beta$ respectively)
suggest a possible way to
quantify an application's dynamism and distribution and thereby find
possible common approaches solutions.  We plan to explore this in our
future work, and also hope others will consider building on these
initial concepts.
